Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-11-12 paper.tex
In depth pass on the first part (section 1 and 2)

\def\journal{Cognitive Science}
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} 
\def\Authors{Julia Fink\,$^{1*}$, S\'{e}verin Lemaignan$^{1}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland 
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

By building upon experimental results from several studies, a comprehensive
synthesis of literature, as well as insights from social psychology,
cognitive science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism. By doing so, we propose a \textbf{formal model} that defines a
\textbf{clear terminology} and accounts for the \textbf{non-monotonic dynamics} of
the phenomenon. We further relate the observed \emph{effects} of anthropomorphism to
their \textbf{underlying cognitive precursors}.

We examine several \textbf{factors} that determine anthropomorphism, namely the
design of the non-human agent, the individual psychological determinants of the
human user, and the context of use. Moreover, we discuss several \textbf{episodes of
interaction} that influence the anthropomorphic projections on the robot, like
the \textbf{novelty effect}, the role of \textbf{familiarization} and the
impact of \textbf{disruptive behaviors}.

Finally, the article also discusses techniques and scenarios to measure anthropomorphism,
and we conclude with a few suggestions on how others can apply our model to
better describe people's tendency to perceive human-like characteristics in non-human
agents.

@@

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction by
considering the phenomenon as a whole. Robotics researchers often tend to
believe that anthropomorphism only describes a static set of human-like features
of a robot (like shape, speech capabilities, facial expressions, etc.). We refer
to these characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that
\emph{anthropomorphism} in general refers to a \emph{social phenomenon} that
emerges from the (real or imagined) \emph{interaction} between a non-human agent
(\eg a robot) and a human~\citep{persson_anthropomorphism_2000}. According
to~\cite{epley_when_2008}, anthropomorphism includes that the human perceives,
for instance, emotional states, motivations, or intentions in the non-human
agent, and tends to ascribe those qualities to it.  As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

To support the reflection on variances and dimensions of anthropomorphism, it may be 
useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study:
                the robot toy box Ranger transports domino tiles between the two
                children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{correlation}
            \caption{\emph{Engagement actions} vs \textit{subjective anthropomorphic
                perception} evidence one cluster of
                high-anthropomorphizers/low-interactors and another group of children who
                interact more with the robot and anthropomorphize it less.}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children's interaction with Ranger, and their anthropomorphic
        projections onto the robot. Overall, large variations of both children's
        behavior and their perception of the robot were found but seemed to be
        correlated to each other.}
    \label{fig:ranger}
\end{figure}

Figure~\ref{fig:qualitative-score} calls for other initial remarks. This diagram
shows data from a (yet unpublished) study in which 13 pairs of children (4-5
years old) played for about 20 minutes a Domino game with a robot toy box (the
robot was tasked to carry domino pieces from one child to the other,
Figure~\ref{fig:ranger-expe}).  The $x$-axis represents the number of social
interactions between the children and the robot (like talking to the robot or
showing objects), based on video annotations. On the $y$-axis, we have a
synthetic qualitative measurement of anthropomorphic projections onto the robot,
based on two interviews with the children. The correlation between both these
values is significantly \emph{negative}: the more the children engaged into
interaction, the less prone to anthropomorphic projections they were. This
diagram also highlights an interesting cluster of children who did not interact
much with the robot, and still reported a strong anthropomorphic perception of
the robot.

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that it supports or improves human-robot interaction?

@@

First, by reviewing what anthropomorphism actually \emph{means}: besides an
\emph{artifact-centered} understanding that explains anthropomorphism from the
human-like design of the non-human artifact, it seems important to explicit the
social component, namely a \emph{human-centered} understanding, of
anthropomorphism.

Second, we propose a model, called the \textit{Dynamics of Anthropomorphism} to
understand how anthropomorphism evolves over time
~\citep{lemaignan2014dynamics}.  These dynamics are non-monotonic, and effects
like the \emph{novelty effect} play here a key role.

Finally, we adopt a third perspective on anthropomorphism by investigating its
\emph{cognitive correlates}. Several studies previously showed that we build
instinctive bonds with robots as soon as we attribute animacy, and we also know
from social psychology some of the mechanisms that lead to
\emph{familiarization}. We attempt here to ``connect the dots'', in order to
build a practical model of the cognitive stages we encounter in human-robot
interaction.

\subsection{Terminology used in this article}

We use the term \textbf{anthropomorphism} to describe the \emph{psychological}
phenomenon, which arises in the \emph{social context} of an interaction between
a human agent and a non-human agent, in our case, a robot (see Figure
\ref{fig:anthropofig}). We use this term to refer to both people's tendency to
\textit{perceive} human-like characteristics, and to the tendency to
\textit{ascribe} or \textit{attribute} such human-like characteristics to the
agent or artifact. Besides, we call {\bf anthropomorphic effects} the
\emph{observable manifestations} of anthropomorphism, \ie the human's behaviors
and actions toward a non-human agent that reflect the underlying cognitive
process of anthropomorphizing.

However, despite the fact that ``anthropomorphism'' literally means ``human
form'' we propose that the term should \textit{not} be used to refer to the
\textit{form / design} of a non-human agent or artifact. To refer to an
imitation of human-like \textit{form / design} of artificial agents, we propose
to use the term \textbf{anthropomorphic design}. This explicit terminology
prevents the commonly encountered confusion between anthropomorphism as a
psychological, subjective phenomenon, and the external, objective factors that
\emph{influence} the phenomenon itself.

In the following, we might use the terms \emph{form} and \emph{design}
interchangeably. Note that the terms \textit{form} or {\it design} do not only
refer to the shape (how it looks like) but to the total expression of the
artifact~\citep{bartneck_shaping_2004}. \citet{disalvo_hug:_2003} argues that
\textit{form} includes the physical shape, materials, and behavioral qualities
of the product. The behavioral aspect is of special importance for interactive
systems that show some degree of autonomy (independence).  Several categories of
anthropomorphic form exist, with no common agreement on the borders and
transitions between those: anthropomorphic, zoomorphic, caricatured, functional
\citep{fong_survey_2003}.

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, we first (in Section \ref{sec:anthropomorphism}) aim
to establish a common ground in the understanding of anthropomorphism. We review
literature from several relevant domains and integrate the different
explanations into an understanding of anthropomorphism as a multi-factor
phenomenon.  Based on this, Section~\ref{sec:our-ideas} proposes a new
conceptual model of the \textit{Dynamics of Anthropomorphism}. This model
describes the variances and dynamics of people's tendency to anthropomorphize an
artificial agent, such as a robot.  Then Section
\ref{sec:cognition-neuroscience} presents a cognitive and neuroscience
perspective on anthropomorphism to consolidate our considerations and the
proposed model.  Section \ref{sec:measuring} looks at how these models can be
operationalized, and in particular, what are the tools at hand to measure
anthropomorphism. Finally, Section \ref{sec:conclusion} concludes this article
and offers starting points for further research on anthropomorphism and the
effect of the human-like design approach in robotics.

@@

Literature on anthropomorphism is quite diverse. There exist different and
partly conflicting definitions of the term anthropomorphism, and it still
appears controversial that humans tend to \textit{``see human
in the non- human''} \citep{epley_seeing_2007}.

As anthropomorphism arises in an interaction between a human and a non-human
agent~\citep{persson_anthropomorphism_2000}, one can also take two main
\textit{perspectives} when trying to explain why humans anthropomorphize~\citep{lee_human_2005}: an artifact-centered perspective and a human-centered
perspective (the grey triangles in Figure \ref{fig:anthropofig}).

In the following we will only briefly present the artifact-centered perspective
to explain anthropomorphism. A review of anthropomorphic design of robots can
also be found in~\cite{fink_anthropomorphism_2012}. Then, we provide a more
detailed review and synthesis of literature that applies a human-centered
perspective to explain anthropomorphism. We aim here at providing an exhaustive
picture of this phenomenon, by summarizing theories from developmental
psychology, cognitive science, neuroscience, and social psychology.

\subsection{Overview of the diverse understandings of anthropomorphism}

Naturally, conflicts in how terms are used lead to a diversity of understandings
of the phenomenon which is described.  One reason for this might be that
anthropomorphism is a phenomenon that is studied in very different domains that
might each integrate their own understanding, use the same term to refer to
different things.  Despite the fact that there is no commonly accepted
definition, the terms \textit{anthropomorphism}, \textit{anthropomorphic} or
\textit{human-like} are often used as if their meanings were clear and agreed
upon \citep{persson_anthropomorphism_2000}. It is even argued that these terms
might be misused \citep{duffy_anthropomorphism_2002,epley_when_2008}.  For
instance, some researchers refer to \textit{``the robot's level of
anthropomorphism''} \citep{bartneck_is_2007,feil-seifer_human-robot_2008},
whereas others disagree to such a statement because they understand that a
system or artifact itself does not \textit{``contain anthropomorphism''} but
only gives rise to the process of anthropomorphizing in a given user and
situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such. We come back on this point in the
next section.

Another paradox is that, while human's tendency to anthropomorphize is commonly
observed, the phenomenon is at the same time still rather poorly understood
\citep{epley_seeing_2007}. The phenomenon itself has been found to be very
complex, sometimes subtle and as such hard to study
\citep{duffy_anthropomorphism_2002,epley_when_2008}. Besides, it is not clear
how to operationalize and measure anthropomorphism (we discuss this question
toward the end of this article, in Section \ref{sec:measuring}).

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{Anthropomorphism is a psychological phenomenon which arises in a
        social context, during an interaction between a human agent and a
        non-human agent. One can take two main perspectives when trying to
        explain anthropomorphism: a human-centered and an artifact-centered
        perspective (grey triangles). Several external factors influence the
        anthropomorphism phenomenon. Person- and artifact-related characteristics,
        but also related to the context of use as well as the actual time spent
        interacting with the robot (interaction history).}

%
%\begin{figure}
%    \centering
%    \includegraphics[width=0.75\columnwidth]{in_our_heads.pdf}
%    \caption{[ALTERNATIVE TO FIGURE \ref{fig:anthropofig}] Anthropomorphism is a
%        subjective, psychological phenomenon which arises in a social context, during an
%        interaction between a human agent and a non-human agent. One can take
%        two main perspectives when trying to explain anthropomorphism: a
%        human-centered and an artifact-centered perspective.
%        Several external factors influence the anthropomorphism phenomenon.
%        Person- and artifact-related characteristics, but also related to the
%        context of use as well as the actual time spent interacting with the
%        robot (interaction history).}
%
%
%    \label{fig:anthropofig-alt}
%\end{figure}
%

@@

The artifact-centered perspective explains anthropomorphism from the
ascertainable parts of the non-human agent, namely the anthropomorphic form of
it. This section deals with the question how far anthropomorphism can be
explained from the design of the artifact.  One of the most prominent working
hypothesis in this field is the Uncanny Valley hypothesis
\citep{mori_uncanny_1970}. In general, it is proposed that human-like designs of
robots can evoke both positive and negative feelings in a human user.

@@

@@

@@

@@

However, we think that this is not the full story. When agreeing to understand
anthropomorphism as a specific type of \textit{experience} that arises in an
\textit{interaction} between a set of user expectations and the external reality
\citep{persson_anthropomorphism_2000}, 
%(see Figure~\ref{fig:anthropomorphism_and_interaction})
it becomes clear that this interaction happens in a specific context, with a
specific history. We think that the context and the history of the interaction
(in Figure \ref{fig:factors} denoted as \textit{time}), are two more factors
that determine anthropomorphism.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{factors}
    \caption{Anthropomorphism is a multi-factor phenomenon. How far people tend
    to anthropomorphize a robot is determined by four main factors: the design
    of the robot $(R)$, the personal characteristics of the human user $(H)$, the
    situation and context in which the interaction occurs $(C)$, and the amount of
    time (repeated interaction) that the user spends to familiarize herself with the
    robot $(t)$.}

    \label{fig:factors}
\end{figure}

Consequently, we propose a holistic understanding of anthropomorphism as a
phenomenon which is determined by multiple factors (see also Figure
\ref{fig:factors}):

@@

In the \emph{Dynamics of Anthropomorphism} model, anthropomorphism is quantified
by a \emph{normalized level of anthropomorphic effects} $\antNorm[R,H,C,t] =
\frac{\ant[R,H,C,t]}{\antMax}$: because anthropomorphic effects are often
qualitative and difficult to quantify on an absolute scale, we present them as a
normalized value, that spans from a minimum (no anthropomorphic effects:
$\ant=0$) to a maximum (noted \AntMax, with $t_{max} = \operatorname{arg\,max}_t
\, \ant[t]$). For the sake of readability, we will use \Ant in the following
sections to refer to the normalized level of anthropomorphic effects. Note also
that we discuss in Section~\ref{sec:measuring} possible methodologies for the
quantitative and qualitative assessment of anthropomorphic effects.

@@

The \textit{Initialization Phase} starts with the so-called \textit{Initial
Capital of Anthropomorphism} ($ICA$), which corresponds to \ICA (value of \Ant
at time point $0$). This value considers the multiple factors presented in
Section~\ref{sec:multiple-factors}, thus it is composed by factors of the human
$H$, the robot $R$, and the context of use $C$. As such, \ICA describes the
value of anthropomorphic effects that can be expected to occur during the first
contact between the human and the robot in a given situation. During this first
phase, pre-conceptions and initial expectations as well as previous experience
of the user with other similar types of robots and similar contexts can have a
strong impact and influence the general tendency to anthropomorphize.

@@

We call \emph{disruptive behaviors} any behavior exhibited by the robot hat
appears unexpected to the user: for instance, a robot may usually follow always
the same route to go from one place in a house to the other, but suddenly it
might change the route. The actual underlying reason may span from a bug to the
detection of a new obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}. We further
discuss the role of disruptive behaviors in cognitive terms in Section
\ref{sec:disruptive}.

>>>>

\def\journal{Cognitive Science}
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} 
\def\Authors{Julia Fink\,$^{1*\dagger}$, S\'{e}verin Lemaignan$^{1\dagger}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland

\vspace{0.5em}\tiny{$\dagger$ These authors have equally contributed to this article.}
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

By building upon experimental results from several studies, a comprehensive
synthesis of literature, as well as insights from social psychology,
cognitive science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism. We propose a \textbf{qualitative model} that defines a
clear terminology and accounts for the \textbf{non-monotonic dynamics} of
the phenomenon. We further relate the observed effects of anthropomorphism to
their \textbf{underlying cognitive precursors}.

We examine several \textbf{factors} that determine anthropomorphism, namely the
design of the non-human agent, the individual psychological determinants of the
human user, and the context of use. Moreover, we discuss several episodes of
interaction that influence the anthropomorphic projections on the robot, like
the \textbf{novelty effect}, the role of \textbf{familiarization} and the
impact of \textbf{disruptive behaviors}.

Finally, the article also discusses techniques to measure anthropomorphism,
and we conclude with a few suggestions on how others can apply our model to
better describe people's tendency to perceive human-like characteristics in non-human
agents.

@@

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
This article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction by
considering the phenomenon as a whole. Robotics researchers often tend to
believe that anthropomorphism only describes a static set of human-like features
of a robot (like shape, speech capabilities, facial expressions, etc.). We refer
to these characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that
\emph{anthropomorphism} in general refers to a \emph{social phenomenon} that
emerges from the (real or imagined) \emph{interaction} between a non-human agent
(a robot in our case) and a human~\citep{persson_anthropomorphism_2000}. According
to~\cite{epley_when_2008}, anthropomorphism includes that the human perceives,
for instance, emotional states, motivations, or intentions in the non-human
agent, and tends to ascribe those qualities to it.  As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

To support the reflection on variances and dimensions of anthropomorphism, it is 
useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the domino study:
                    the robot toy box \emph{Ranger} transports domino tiles between the two
                children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{correlation}
            \caption{\emph{Actions reflecting engagement} vs \textit{subjective anthropomorphic
                perception} evidence one cluster of
                high-anthropomorphizers/low-interactors and another group of children who
                interact more with the robot and anthropomorphize it less.}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Investigating children's interactions with a robot in a playful
            context: anthropomorphic perception does not necessarily positively
            correlate with engagement.}
    \label{fig:ranger}
\end{figure}

Figure~\ref{fig:qualitative-score} calls for other initial remarks. This diagram
shows data from a (yet unpublished) study in which 13 pairs of children (4-5
years old) played for about 20 minutes a domino game with a robot toy box (the
robot was tasked to carry domino pieces from one child to the other,
Figure~\ref{fig:ranger-expe}).  The $x$-axis represents the number of social
interactions between the children and the robot (like talking to the robot or
showing objects), based on video annotations. On the $y$-axis, we have a
synthetic qualitative measurement of anthropomorphic perception of the the
robot, based on two interviews with the children. The scatter plot evidences one
cluster of high-anthropomorphizers/low-interactors and a another group of
children who interact more with the robot and anthropomorphize it less: the
more the children engaged into interaction, the less prone to an anthropomorphic
perception they were.

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that it supports or improves the design of sustainable human-robot interaction?

@@

First, by reviewing what anthropomorphism actually means: besides an
\emph{artifact-centered} understanding that explains anthropomorphism from the
human-like design of the non-human artifact, it seems important to explicit the
social component, namely a \emph{human-centered} understanding, of
anthropomorphism.

Second, we propose a qualitative model of the \textit{dynamics of
anthropomorphism} to understand how anthropomorphism evolves over time. These
dynamics are typically non-monotonic, and effects like the \emph{novelty effect} play here
a key role.

Finally, we synthetize and integrate in an explanatory model the \emph{cognitive
mechanisms} that underlie anthropomorphism. Several studies previously showed
that we build instinctive bonds with robots as soon as we attribute animacy, and
we also know from social psychology some of the mechanisms that lead to
\emph{familiarization}. We attempt here to build a practical model of the
cognitive stages we encounter in human-robot interaction.

\paragraph{Terminology used in this article\\}

While we discuss in section~\ref{meaning} the different meanings of
anthropomorphism found in literature, it seems appropriate to already precisely
define the terminology used in this article.

We use the term \textbf{anthropomorphism} to describe the \emph{psychological}
phenomenon which arises in the \emph{social context} of an interaction between a
human agent and a non-human agent, in our case a robot (see Figure
\ref{fig:anthropofig}). We use this term to refer to both people's tendency to
\textit{perceive} (\textbf{anthropomorphic perceptions}) human-like
characteristics, and to the tendency to \textit{ascribe} or \textit{attribute}
such human-like characteristics to the agent or artifact
(\textbf{anthropomorphic projections}). Besides, we call {\bf anthropomorphic
effects} the \emph{observable manifestations} of anthropomorphism, \ie the
human's behaviors and actions toward a non-human agent that reflect the
underlying cognitive process of anthropomorphizing.

Despite the fact that ``anthropomorphism'' literally means ``human
form'' we propose that the term should \textit{not} be used to refer to the
\textit{form / design} of a non-human agent or artifact. To refer to an
imitation of human-like \textit{form / design} of artificial agents, we propose
to use the term \textbf{anthropomorphic design}. This explicit terminology
prevents the commonly encountered confusion between anthropomorphism as a
psychological, subjective phenomenon, and the external, objective factors that
\emph{influence} the phenomenon itself.

In the following, we might use the terms \emph{form} and \emph{design}
interchangeably. Note that the terms \textit{form} or {\it design} do not only
refer to the shape (how it looks like) but to the total expression of the
artifact~\citep{bartneck_shaping_2004}. \citet{disalvo_hug:_2003} argues that
\textit{form} includes the physical shape, materials, and behavioral qualities
of the product. The behavioral aspect is of special importance for interactive
systems that show some degree of autonomy (independence).

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, section \ref{sec:anthropomorphism} establishes a
common ground in the understanding of anthropomorphism. We review literature
from several relevant domains and integrate the different explanations into an
understanding of anthropomorphism as a multi-factor phenomenon.  Based on this,
section~\ref{sec:our-ideas} proposes a new qualitative model of the
\textit{dynamics of anthropomorphism}. This model describes the variances and
dynamics of people's tendency to anthropomorphize an artificial agent, such as a
robot. Section \ref{sec:cognition-neuroscience} then investigates the cognitive
and neuro-biological mechanisms of anthropomorphism to consolidate our
considerations and the proposed model.  Section \ref{sec:measuring} looks at how
these models can be operationalized, and in particular, what are the tools at
hand to measure anthropomorphism. Finally, section \ref{sec:conclusion}
concludes this article and offers starting points for further research on
anthropomorphism and the effect of the human-like design approach in robotics.

@@

Literature on anthropomorphism is quite diverse. There exist different and
partly conflicting definitions of the term anthropomorphism, and it still
appears controversial that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. As anthropomorphism arises in an interaction
between a human and a non-human agent~\citep{persson_anthropomorphism_2000}, one
can also take two main \textit{perspectives} when trying to explain why humans
anthropomorphize~\citep{lee_human_2005}: an artifact-centered perspective and a
human-centered perspective.

After discussing and clarifying the different meanings of anthropomorphism, we
present the main aspects of an \emph{artifact-centered} perspective on
anthropomorphism. We then provide a synthesis of literature that study
 anthropomorphism from the perspective of the human: we aim here at providing a broad
picture of this phenomenon by summarizing theories from developmental
psychology, cognitive science, neuroscience, and social psychology.

We then briefly discuss the impact of the interaction context on
anthropomorphism, before restating anthropomorphism as a \emph{multi-factor
phenomenon}.

\subsection{Overview of the diverse understandings of anthropomorphism}
\label{meaning}

Naturally, conflicts in how terms are used lead to a diversity of understandings
of the phenomenon which is described: anthropomorphism is a phenomenon that is
studied in very different domains that each integrate their own understanding,
use the same term to refer to different things.  Despite the fact that there is
no commonly accepted definition, the terms \textit{anthropomorphism},
\textit{anthropomorphic} or \textit{human-like} are often used as if their
meanings were clear and agreed upon \citep{persson_anthropomorphism_2000}. It is
however argued that these terms might be misused
\citep{duffy_anthropomorphism_2002,epley_when_2008}.  For instance, some
researchers refer to \textit{``the robot's level of anthropomorphism''}
\citep{bartneck_is_2007,feil-seifer_human-robot_2008}, whereas others disagree
to such a statement because they understand that a system or artifact itself
does not \textit{``contain anthropomorphism''} (anthropomorphism is not
intrinsic to the robot) but only gives rise to the process of anthropomorphizing
in a given user and situation. Consequently,
\cite{persson_anthropomorphism_2000} conclude that anthropomorphism emerges in
the \textit{interaction} between the technology and the user, and has to be
understood as such. We come back on this point in the next section.

Another paradox is that, while human's tendency to anthropomorphize is commonly
observed, the phenomenon is at the same time still rather poorly understood
\citep{epley_seeing_2007}. The phenomenon itself has been found to be very
complex, sometimes subtle and as such hard to study
\citep{duffy_anthropomorphism_2002,epley_when_2008}. Besides, it is not clear
how to operationalize and measure anthropomorphism (we discuss this specific question
in section \ref{sec:measuring}).

@@

@@

The \emph{artifact-centered} perspective explains anthropomorphism from the
ascertainable parts of the non-human agent, namely the anthropomorphic form of
it. This section deals with the question how far anthropomorphism can be
explained from the design of the artifact.  One of the most prominent working
hypothesis in this field is the Uncanny Valley hypothesis
\citep{mori_uncanny_1970}. In general, it is proposed that human-like designs of
robots can evoke both positive and negative feelings in a human user.

@@

\paragraph{The Physical shape} of a robot refers to `how the robot looks like'
and includes its general form, and other visible features, such as its size, the
materials used, its color, and specific details. Generally, robots can be
divided into human-like (anthropomorphic) and mechanical (non-anthropomorphic)
robots. However, it makes sense to be slightly more specific about the form, and
\cite{fong_survey_2003} suggest four categories: anthropomorphic, zoomorphic,
caricatured, functional. The borders and transitions between these categories
are not very well defined.  Of particular importance seems to be the presence or
absence of specific human body parts, such as arms and hands, legs, a torso or a
head and face. Of those, the faces of robots have received remarkable attention
concerning how they are perceived and accepted. \cite{disalvo_all_2002} studied
what features and dimensions of a humanoid robot's face most dramatically
contribute to people's perception of its humanness. The authors analyzed 48
humanoid robot heads and found that the presence of certain features, the
dimensions of the head, and the total number of facial features heavily
influence the perception of humanness in robot heads.

\paragraph{Expression, communication, and behavior} refers to how the robot
expresses its human qualities through interacting and behaving. In fact, many
humanoid robots express their human-likeness not in physical shape but only
through the \textit{suggestion} of features~\citep{disalvo_all_2002}. Besides
this, communication and interaction modalities of a robot are likely to play a
role to what extent the robot will be anthropomorphized. Can the user interact
with the robot through a touch interface, or using speech? Does the robot
respond using linguistic utterances or non-linguistic sounds? Is the voice
human-like or synthetic, on-board or coming from an external second device? A
study by~\cite{takayama_im_2009} found that this has an effect on people's
perception of trust in the robot and how convincing the robot appears to them.
Besides this, we expect that a robot's behavioral aspects concerning
human-social interaction have an impact. Does the robot use polite formulations,
etiquette, appropriate facial expressions? Are its responses and reactions
timely? Usually, robots that directly interact with humans are slow, which
appears not very human-like to most participants.

\paragraph{The functionality, and purpose} are a robot-related characteristic
that also partly determines the context of use. This aspect includes what kind
of expectations the user may have of the robot, and how the user imagines using
the robot. Is there only one user or a group of people interacting with the
robot at the same time? Is the purpose of the robot a playful or a serious one?
Does the robot have a purpose at all~\citep{kaplan_free_2000}? We further
discuss this aspect in the section related to the context of use.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
% Context
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Role of the Context}

\textit{Context of use} is defined by the ISO as \textit{``users, tasks,
equipment (hardware, software and materials), and the physical and social
environments in which a product is used''}. Accordingly, there are various
aspects that can impact how a robot is experienced. The real or imagined
\textit{purpose} of a robot, including the application context in which it is
typically used (\eg office environment, home, rehabilitation center, public
space, \etc) as well as the task context (\eg serious or playful task;
composition of human-robot team; number of users) and the (social) role in which
the robot is used / experienced are taken into account. Generally, when the
context of use of a robot is social, entertaining or playful, it will enhance
anthropomorphism, compared to when the context is a routine or focused serious
task (security, rescue, \etc). \cite{joosse_what_2013} showed, for instance,
that when the same robot ({\sc nao}) was used in a different task context (cleaning
task \vs tour guide), users ascribed different personalities to the robot. \cite{goetz_cooperation_2002}
revealed a link between a robot's context of use and people's perception of the
robot. The authors found that people prefer a serious robot for serious tasks
and a less serious robot for more playful tasks
. \cite{kaplan_free_2000}
discussed the role of uselessness in the design of robots. The author argues
that artificial pets like AIBO have no real purpose, in a sense that they do not
provide any kind of service or, in other words, \textit{``they are not doing
what you tell them to do''} and \textit{``they might refuse the order of its
owner''}. It may be this very aspect that increases people's tendency to
anthropomorphize the robot. According to Kaplan, it is in our daily use language
that we tend to attribute intentions to devices that are not doing their job
well.

The characteristics of the context can also impact on the user experience and in
turn foster or hinder anthropomorphism indirectly. For instance, when there are
several friends of the main user interacting simultaneously with the robot, the
likelihood of anthropomorphism may be increased. The context here is already
social when several people who know each other engage in social interaction
among themselves, so the robot may be perceived as being somewhat part of the
social interaction, and hence it may be attributed human-social qualities
\citep{baxter2013do}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{Anthropomorphism is a multi-factor phenomenon. How far people tend
    to anthropomorphize a robot is determined by four main factors: the design
    of the robot $(R)$, the personal characteristics of the human user $(H)$, the
    situation and context in which the interaction occurs $(C)$, and the amount of
    time (repeated interaction) that the user spends to familiarize herself with the
    robot $(t)$.}

@@

However, we think that this is not the full story. When agreeing to understand
anthropomorphism as a specific type of \textit{experience} that arises in an
\textit{interaction} between a set of user expectations and the external reality
\citep{persson_anthropomorphism_2000}, 
%(see Figure~\ref{fig:anthropomorphism_and_interaction})
it becomes clear that this interaction happens in a specific context, with a
specific history. We think that the context and the history of the interaction
(in Figure \ref{fig:anthropofig} denoted as \textit{time}), are two more factors
that determine anthropomorphism.

Consequently, we propose a holistic understanding of anthropomorphism as a
phenomenon which is determined by multiple factors (see also Figure
\ref{fig:anthropofig}):

@@

In the \emph{Dynamics of Anthropomorphism} model, anthropomorphism is quantified
by a \emph{normalized level of anthropomorphic effects} $\antNorm[R,H,C,t] =
\frac{\ant[R,H,C,t]}{\antMax}$: because anthropomorphic effects are often
qualitative and difficult to quantify on an absolute scale, we present them as a
normalized value, that spans from a minimum (no anthropomorphic effects:
$\ant=0$) to a maximum (noted \AntMax, with $t_{max} = \operatorname{arg\,max}_t
\, \ant[t]$). For the sake of readability, we will use \Ant in the following
sections to refer to the normalized level of anthropomorphic effects. Note also
that we discuss in section~\ref{sec:measuring} possible methodologies for the
quantitative and qualitative assessment of anthropomorphic effects.

@@

The \textit{Initialization Phase} starts with the so-called \textit{Initial
Capital of Anthropomorphism} ($ICA$), which corresponds to \ICA (value of \Ant
at time point $0$). This value considers the multiple factors presented in
section~\ref{sec:multiple-factors}, thus it is composed by factors of the human
$H$, the robot $R$, and the context of use $C$. As such, \ICA describes the
value of anthropomorphic effects that can be expected to occur during the first
contact between the human and the robot in a given situation. During this first
phase, pre-conceptions and initial expectations as well as previous experience
of the user with other similar types of robots and similar contexts can have a
strong impact and influence the general tendency to anthropomorphize.

@@

We call \emph{disruptive behaviors} any behavior exhibited by the robot hat
appears unexpected to the user: for instance, a robot may usually follow always
the same route to go from one place in a house to the other, but suddenly it
might change the route. The actual underlying reason may span from a bug to the
detection of a new obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}. We further
discuss the role of disruptive behaviors in cognitive terms in section
\ref{sec:disruptive}.
